
\subsection{Publications}
\begin{frame}
	\frametitle{Publications}
	 	\begin{itemize}
	 		\tiny \item \textbf{Olalekan Ogunmolu}, A Multi-DOF Soft Robot Mechanism for Patient Motion Correction and Beam Orientation Selection	in Cancer Radiation Therapy. PhD Thesis Manuscript. Hyperlink: \href{https://ecs.utdallas.edu/~opo140030/media/Papers/thesis.pdf}{\textcolor{blue}{utdallas.edu/~opo140030/media/Papers/thesis.pdf}}
	 		%
	 		\vspace{0.5em}
	 		%
	 		\tiny \item Azar Sadeghnejad Barkousaraie, \textbf{Olalekan Ogunmolu}, Steve Jiang, and Dan Nguyen.  \href{}{\bl{A Fast Deep Learning Approach for Beam Orientation Selection Using Supervised Learning with Column Generation on IMRT Prostate Cancer Patients}}. \href{}{Under review at \emph{Medical Physics}} (Journal), April 2019.
	 		%
	 		\vspace{0.5em}
	 		%
	 		\tiny \item \textbf{Olalekan Ogunmolu}, Michael Folkerts, Dan Nguyen, Nicholas Gans, and Steve Jiang.  \href{}{\bl{Deep BOO! Automating Beam Orientation Selection in Intensity Modulated Radiation Therapy}}. \textit{\bl{\href{http://algorithmic-robotics.org/wafr2018/}{Algorithmic Foundations of Robotics XIII, International Workshop}}} (\emph{WAFR}), Mérida, Mexico. December 2018. Published in Springer's Proceedings in Advanced Robotics (SPAR) Book. 2019.
	 		%
	 		\vspace{0.5em}
	 		%
	 		\tiny \item \textbf{Olalekan Ogunmolu}, Nicholas Gans, Tyler Summers. \href{}{\bl{Minimax Iterative Dynamic Game: Application to Nonlinear Robot Control Tasks}}. \textit{IEEE/RSJ International Conference on Intelligent Robots and Systems} (\emph{IROS}), Madrid, Spain. October 2018.
	 		%
	 		\vspace{0.5em}
	 		%		
	 		\tiny \item \textbf{Olalekan Ogunmolu}, Adwait Kulkarn, Yonas Tadesse, Xuejun Gu, Steve Jiang, and Nicholas Gans. \href{https://arxiv.org/abs/1703.03821}{\bl{Soft-NeuroAdapt: A 3-DOF Neuro-Adaptive Pose Correction System For Frameless and Maskless Cancer Radiotherapy}}. \textit{IEEE/RSJ International Conference on Intelligent Robots and Systems} (\emph{IROS}), Vancouver, BC, Canada. September 2017. DOI: 10.1109/IROS.2017.8206211.
	 		%
	 		\vspace{0.5em}
	 		%
	 		\tiny \item Azar Sadeghnejad Barkousaraie,  \textbf{Olalekan Ogunmolu}, Steve Jiang, and Dan Nguyen.  \href{}{\bl{Using supervised learning and guided Monte Carlo tree search for beam orientation optimization in radiation therapy}}. Under review at \textit{\bl{\href{http://iccr-mcma.org/iccr2019/}{International Conference on Medical Image Computing and Computer Assisted Intervention, XXII}}} (\emph{MICCAI}), Shenzhen, China. October 2019.
	 	\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Publications}
	\begin{itemize}
		\tiny \item Azar Sadeghnejad Barkousaraie,  \textbf{Olalekan Ogunmolu}, Steve Jiang, and Dan Nguyen.  \href{}{\bl{Deep Learning Neural Network for Beam Orientation Optimization}}. To appear in \textit{\bl{\href{http://iccr-mcma.org/iccr2019/}{International Conference on the use of Computers in Radiation Therapy XVI}}} (\emph{ICCR}), Montreal, CA. June 2019.
		%
		\vspace{0.5em}
		%	
		\tiny \item \textbf{Olalekan Ogunmolu}, Dan Nguyen, Xun Jia, Weiguo Lu, Nicholas Gans, and Steve Jiang. \href{}{\bl{Automating Beam Orientation Optimization for IMRT Treatment Planning: A Deep Reinforcement Learning Approach}}.  	%\newline 
		Selected for Oral Presentation at the \textit{John R. Cameron Young Investigators Symposium -- 60th Annual Meeting of the American Association of Physicists in Medicine}, Nashville, TN (\emph{AAPM}). July 2018.
		%
		\vspace{0.5em}
		%
		\tiny \item Yara Almubarak, Joshi Aniket, \textbf{Olalekan Ogunmolu}, Xuejun Gu, Steve Jiang, Nicholas Gans, and Yonas Tadesse, \bl{Design and Development of Soft Robots for Head and Neck Cancer Radiotherapy.} \emph{\href{http://spie.org/conferences-and-exhibitions/smart-structures/nde?SSO=1}{\bl{SPIE: Smart Structures + Nondestructive Evaluation}},} (\emph{SPIE}), Denver, CO, U.S.A. March 2018.
		%
		\vspace{0.5em}
		%
		\tiny \item \textbf{Olalekan Ogunmolu}, Xuejun Gu, Steve Jiang, and Nicholas Gans. \href{http://ecs.utdallas.edu/~opo140030/media/Papers/CASE2016.pdf}{\bl{Vision-based control of a soft-robot for Maskless Cancer Radiotherapy}}. \emph{IEEE Conference on Automation Science and Engineering (CASE)}, Fort-Worth, Texas, August 2016. DOI: 10.1109/CoASE.2016.7743378.
		%
		\vspace{0.5em}
		%
		\tiny \item \textbf{Olalekan Ogunmolu}, Xuejun Gu, Steve Jiang, and Nicholas Gans. \href{http://ecs.utdallas.edu/~opo140030/media/Papers/CASE2015/CASE2015.pdf}{\bl{A Real-Time Soft-Robotic Patient Positioning System for Maskless Head-and-Neck Cancer Radiotherapy}}. \emph{IEEE Conference on Automation Science and Engineering (CASE)}, Gothenburg, Sweden, August 2015. DOI: 10.1109/CoASE.2015.7294318.
	\end{itemize}
\end{frame}