
\begin{abstract}
	Finding the optimal beam angle set from a robot's tool frame from which to aim radiation to cancer tumor(s) whilst simultaneously minimizing radiation intensity to organs-at-risks during intensity modulated radiation therapy is an exhaustive combinatorial search/human-based parameters-tuning process. So far, traditional optimization approaches for solving this problem have not yielded real-time feasible results. Aiming to automate the selection process, we introduce novel techniques from (i) pattern recognition, (ii) monte carlo evaluations, (iii) game theory, and (iv)  neuro-dynamic programming. This is so as to find admissible policies for beamlets selection duringh IMRT treatment planning. In this sentiment, we optimize a deep neural network policy which guides monte carlo simulations and evaluations of promising beamlets within the robot gantry's phase space. Our goal is not to exclusively find the minimum of an objective as traditional approaches try to do, but rather to go uphill every so often so as to arrive at a pair of saddle point equilibrium strategy for the system. We let  two fictitious neural network players,  within a zero-sum markov game framework  alternatingly play a best response to their opponent's mixed strategy profile during epsidoes of a markov decision process. At inference time, we run the trained policy on test patients' target volumes to predict feasible beam angles. This part of our paper describes the problem and formally introduces our approach. The implementation  details and results are presented in the addendum to this paper~\cite{Ogunmolu2018b}.
\end{abstract}
